{\Huge Supersymmetry} \\ \vspace{5mm} Lecture notes for FYS5190/FYS9190 Supersymmetry

### 2025-10-20 lagrangian.tex
Fixes for lecture 21.10.25

The requirement of renormalisability of the resulting quantum field theory puts further restrictions on the fields in $\mathcal{L}$. We can at most have three factors of scalar superfields in each term of the superpotential, and two factors of scalar superfields in the Kähler potentia,  for details see {\it e.g.}\ Wess \& Bagger \cite{Bagger:1983mv}.\footnote{The argument is similar to ordinary quantum field theory, terms with couplings with negative mass dimension is forbidden.} 
Since the total action must be real, the (almost) most general supersymmetry Lagrangian that can be written in terms of scalar superfields is:
\[\mathcal{L} = \Phi^\dagger_i\Phi_i + \bar{\theta}\bar{\theta}W[\Phi_i] + \theta\theta W[\Phi^\dagger_i].\]
Here the first term is called the {\bf kinetic term},\footnote{The constant in front of this term can always be chosen to be one because we can rescale the whole Lagrangian. It is possible to start from more general Kähler potentials, but this is mostly beyond the scope of these notes, although we will return to modify this term to account for gauge invariance.} and $W$ is the symbol for the {\bf superpotential}, which by renormalisability is restricted to
\begin{equation}
W[\Phi] = g_i \Phi_i + m_{ij}\Phi_i \Phi_j + \lambda_{ijk} \Phi_i \Phi_j \Phi_k.
\end{equation}
This means that to specify a supersymmetric Lagrangian we only need to specify the superfield content of the model and the form of the superpotential.

Dimension counting, starting from $[\mathcal{L}]=M^2$ and $[\Phi]=1$, means that the superpotential terms must each be $M^3$, for the couplings this gives $[g_i]=M^2$, $[m_{ij}]=M$ and $[\lambda_{ijk}]=1$. Notice also that $m_{ij}$ and $\lambda_{ijk}$ must be symmetric since the superfields commute.

@@

Let us first look at the transformation of the superpotential $W$ under the gauge transformation of the scalar superfields in (\ref{eq:scalar_supergauge}):
\[W[\Phi] \to W[\Phi'] = g_i e^{iq_i\Lambda }\Phi_i + m_{ij}e^{i(q_i+q_j)\Lambda}\Phi_i\Phi_j + \lambda_{ijk}e^{i(q_i+q_j+q_k)\Lambda}\Phi_i\Phi_j\Phi_k.\]
Requiring gauge invariance, $W[\Phi] = W[\Phi']$, we must have:
\begin{eqnarray}
q_i &=& 0 ~~\text{or}~g_i = 0, \\
q_i+q_j &=& 0 ~~\text{or}~~m_{ij} = 0,  \\
q_i+q_j+q_k &=& 0  ~~\text{or}~~ \lambda_{ijk}=0.
\end{eqnarray}
This puts great restrictions on the form of the superpotential and the charge assignments of the superfields, as in ordinary gauge theories.

@@

%%%%%%%%%%%%%%%%%%%
\section{Non-Abelian gauge theories}
%%%%%%%%%%%%%%%%%%%
How do we extend the above to deal with the much more complicated non-abelian gauge theories? Let us take a gauge group $G$ with the Lie algebra of group generators $t_a$ 
\begin{equation}
[t_a, t_b] = if_{ab}{}^c t_c,\label{eq:Liealgebrawta}
\end{equation} 
where $f_{ab}{}^{c}$ are the structure constants. We recap that for an element $g$ in the group $G$ we want to write down a unitary representation $U(g)$ that transforms a vector $\Psi$  in the representation by $\Psi \to \Psi' = U(g)\Psi$.\footnote{Of, course, you may ask, how do we even know that we can find a unitary representation for a particular Lie group? For the unitary group this is by definition, but is it possible in general? It turns out that this is always true for {\bf compact Lie groups} which includes many of the matrix groups $U(n)$, $SU(n)$, $O(n)$, $SO(n)$, $Sp(2n)$, and $Sp(n)$.}  
With an exponential map we can write this representation as $U(g) = e^{i\lambda^a t_a}$, where $\lambda^a$ are the parameters and $t_a$ are hermitian operators, as you may perhaps have expected.

As in ordinary gauge theories, we simply copy the structure of the abelian (super)gauge transformation, and transform (a vector of) scalar superfields $\Phi$ under a non-abelian group as
\[\Phi \to \Phi'=e^{iq \Lambda^aT_a}\Phi,\]
in a non-abelian supergauge transformation. Here $q$ is the charge of $\Phi$ under $G$. At this point we need to choose a particular representation, reflected in a particular choice for the generators which we write as $T_a$. Since for gauge groups we are almost exclusively interested in groups defined by a matrix representation, $U(g)$ will be a matrix with dimension fixed by the dimension chosen for the representation. Again we can easily show that we must require that the $\Lambda^a$ has the defining property of a left-handed scalar superfield for $\Phi$ to transform to a left-handed scalar superfield.

@@

We also want a similar construction for the kinetic terms as for abelian gauge theories, $\Phi^\dagger e^{qV^aT_a}\Phi$, to be invariant under non-abelian gauge transformations. Now, using that the generators $T_a$ are hermitian, 
\[\Phi^\dagger e^{qV^aT_a}\Phi \to \Phi'^\dagger e^{qV'^aT_a}\Phi' = \Phi^\dagger e^{-iq\Lambda^{a\dagger} T_a} e^{qV'^aT_a}e^{iq\Lambda^aT_a}\Phi,\]
so, in order to have gauge invariance, we have to require that the vector superfield $V$ transforms as:\footnote{This is independent of our choice of representation for the gauge group for the supergauge transformation.}
\begin{equation}
e^{qV'{}^aT_a} = e^{iq\Lambda^a{}^\dagger T_a}e^{qV^aT_a} e^{-iq\Lambda^a T_a}.
\end{equation}

@@

If we expand $W_A$ in the component fields we find, as we might have hoped, that it contains the ordinary field strength tensor:
\[F_{\mu\nu}^a = \partial_\mu V^a_\nu - \partial_\nu V^{a}_{\mu} + qf_{bc}{}^aV_\mu^bV_\mu^c\]
and that the trace indeed contains terms with $F_{\mu\nu}^{a}F^{\mu\nu a}$.

@@

In Sec.~\ref{sec:superalgebra} we claimed there there is no nontrivial interaction between a gauge group and the supersymmetry generators for $N=1$ supersymmetry. This is not entirely true as we may have non-trivial commutators for the generator $R$ of an abelian group, {\it i.e.}
\begin{equation}
[Q_A, R] = Q_A, \quad  [\bar Q_{\dot A}, R ] = -\bar Q_{\dot A}.
\end{equation}
Using this we define a $U(1)_R$ transformation called {\bf $\mathbf R$-symmetry} as
\begin{equation}
\theta_A \to \theta_A' = e^{i\alpha} \theta_A, \quad \bar\theta_{\dot A} \to  \bar\theta'_{\dot A}= e^{-i\alpha} \bar\theta_{\dot A},
\end{equation}
where $\alpha$ is the parameter of the transformation and the charge of $\theta$ and $\bar\theta$ under the transformation is 1 and $-1$, respectively.
If a Lagrangian is to be invariant under such a transformation then a superfield $\Phi$ must transform as
\begin{equation}
\Phi(x,\theta,\bar\theta)\to \Phi'(x,\theta,\bar\theta)= e^{ir_\Phi\alpha}\Phi(x,e^{-i\alpha} \theta,e^{i\alpha} \bar\theta),
\end{equation}
where $r_\Phi$ is the charge of that superfield under the transformation. The charge of a product of two superfields is then just the sum of the charges of the fields. This is required so that the kinetic term in the Lagrangian is invariant under the transformation.

This means that for a scalar (left- or right-handed) superfield the component fields $A$, $\psi$, and $F$ must transform as
\begin{equation}
A\to e^{ir_\Phi\alpha}A,\quad \psi\to e^{i(r_\Phi-1)\alpha} \psi,\quad F\to e^{i(r_\Phi-2)\alpha} F,
\end{equation}
while vector superfields must have zero $R$-charge since they are real, and their component fields, in the Wess-Zumino gauge must thus transform as
\begin{equation}
V_\mu \to V_\mu ,\quad \lambda \to e^{i\alpha} \lambda,\quad D\to D.
\end{equation}
We can check that all the terms in the Lagrangian (\ref{eq:SUSY_Lagrangian}) are invariant under this transformation, with the exception of the superpotential $W$, which must have total $R$-charge 2. This will exclude most allowed charge assignments for the the scalar superfields, and in the minimal supersymmetric models we shall look at there will be no viable continuous $R$-symmetry. However, $R$-symmetry has importance in building models of supersymmetry breaking, and could remain as a broken symmetry in the model.

>>>>

The requirement of renormalisability of the resulting quantum field theory puts further restrictions on the fields in $\mathcal{L}$. We can at most have three factors of scalar superfields in each term of the superpotential, and two factors of scalar superfields in the Kähler potentia,  for details see {\it e.g.}\ Wess \& Bagger \cite{Bagger:1983mv}.\footnote{The argument is similar to ordinary quantum field theory, terms with couplings with negative mass dimension is forbidden.} 
Since the total action must be real, the (almost) most general supersymmetry Lagrangian that can be written in terms of scalar superfields is:
\[\mathcal{L} = \Phi^\dagger_i\Phi_i + \bar{\theta}\bar{\theta}W[\Phi_i] + \theta\theta W[\Phi^\dagger_i].\]
Here the first term is called the {\bf kinetic term},\footnote{The constant in front of this term can always be chosen to be one because we can rescale the whole Lagrangian. It is possible to start from more general Kähler potentials, but this is mostly beyond the scope of these notes, although we will return to modify this term to account for gauge invariance.} and $W$ is the symbol for the {\bf superpotential}, which by renormalisability is restricted to
\begin{equation}
W[\Phi] = t_i \Phi_i + m_{ij}\Phi_i \Phi_j + \lambda_{ijk} \Phi_i \Phi_j \Phi_k.
\end{equation}
This means that to specify a supersymmetric Lagrangian we only need to specify the superfield content of the model and the form of the superpotential.

Dimension counting, starting from $[\mathcal{L}]=M^2$ and $[\Phi]=1$, means that the superpotential terms must each be $M^3$, for the couplings this gives $[t_i]=M^2$, $[m_{ij}]=M$ and $[\lambda_{ijk}]=1$. Notice also that $m_{ij}$ and $\lambda_{ijk}$ must be symmetric since the superfields commute.

@@

Let us first look at the transformation of the superpotential $W$ under the gauge transformation of the scalar superfields in (\ref{eq:scalar_supergauge}):
\[W[\Phi] \to W[\Phi'] = t_i e^{iq_i\Lambda }\Phi_i + m_{ij}e^{i(q_i+q_j)\Lambda}\Phi_i\Phi_j + \lambda_{ijk}e^{i(q_i+q_j+q_k)\Lambda}\Phi_i\Phi_j\Phi_k.\]
Requiring gauge invariance, $W[\Phi] = W[\Phi']$, we must have:
\begin{eqnarray}
q_i &=& 0 ~~\text{or}~t_i = 0, \\
q_i+q_j &=& 0 ~~\text{or}~~m_{ij} = 0,  \\
q_i+q_j+q_k &=& 0  ~~\text{or}~~ \lambda_{ijk}=0.
\end{eqnarray}
This puts great restrictions on the form of the superpotential and the charge assignments of the superfields, as in ordinary gauge theories.

@@

%%%%%%%%%%%%%%%%%%%
\section{Non-Abelian gauge theories}
%%%%%%%%%%%%%%%%%%%
How do we extend the above to deal with the much more complicated non-abelian gauge theories? Let us take a gauge group $G$ with the Lie algebra of group generators $t_a$ 
\begin{equation}
[t_a, t_b] = if_{ab}{}^c t_c,\label{eq:Liealgebrawta}
\end{equation} 
where $f_{ab}{}^{c}$ are the structure constants. We recap that for an element $g$ in the group $G$ we want to write down a unitary representation $U(g)$ that transforms a vector $\boldsymbol\Psi$  in the representation by $\boldsymbol\Psi \to \boldsymbol\Psi' = U(g)\boldsymbol\Psi$.\footnote{Of, course, you may ask, how do we even know that we can find a unitary representation for a particular Lie group? For the unitary group this is by definition, but is it possible in general? It turns out that this is always true for {\bf compact Lie groups} which includes many of the matrix groups $U(n)$, $SU(n)$, $O(n)$, $SO(n)$, $Sp(2n)$, and $Sp(n)$.}  
With an exponential map we can write this representation as $U(g) = e^{i\lambda^a t_a}$, where $\lambda^a$ are the parameters and $t_a$ are hermitian operators, as you may perhaps have expected.

As in ordinary gauge theories, we simply copy the structure of the abelian (super)gauge transformation, and transform (a vector of) scalar superfields $\boldsymbol\Phi$ under a non-abelian group as
\[\boldsymbol\Phi \to \boldsymbol\Phi'=e^{iq \Lambda^aT_a}\boldsymbol\Phi,\]
in a non-abelian supergauge transformation. Here $q$ is the charge of $\boldsymbol\Phi$ under $G$. At this point we need to choose a particular representation, reflected in a particular choice for the generators which we write as $T_a$. Since for gauge groups we are almost exclusively interested in groups defined by a matrix representation, $U(g)$ will be a matrix with dimension fixed by the dimension chosen for the representation. Again we can easily show that we must require that the $\Lambda^a$ has the defining property of a left-handed scalar superfield for $\boldsymbol\Phi$ to transform to a left-handed scalar superfield.

@@

We also want a similar construction for the kinetic terms as for abelian gauge theories, $\boldsymbol\Phi^\dagger e^{qV^aT_a}\boldsymbol\Phi$, to be invariant under non-abelian gauge transformations. Now, using that the generators $T_a$ are hermitian, 
\[\boldsymbol\Phi^\dagger e^{qV^aT_a}\boldsymbol\Phi \to \boldsymbol\Phi'^\dagger e^{qV'^aT_a}\boldsymbol\Phi' = \boldsymbol\Phi^\dagger e^{-iq\Lambda^{a\dagger} T_a} e^{qV'^aT_a}e^{iq\Lambda^aT_a}\boldsymbol\Phi,\]
so, in order to have gauge invariance, we have to require that the vector superfields $V^a$ transform as:\footnote{This is independent of our choice of representation for the gauge group for the supergauge transformation.}
\begin{equation}
e^{qV'{}^aT_a} = e^{iq\Lambda^a{}^\dagger T_a}e^{qV^aT_a} e^{-iq\Lambda^a T_a}.
\end{equation}

@@

If we expand $W_A$ in the component fields we find, as we might have hoped, that it contains the ordinary field strength tensor:
\[F_{\mu\nu}^a = \partial_\mu V^a_\nu - \partial_\nu V^{a}_{\mu} + qf_{bc}{}^aV_\mu^bV_\nu^c\]
and that the trace indeed contains terms with $F_{\mu\nu}^{a}F^{\mu\nu a}$.

@@

In Sec.~\ref{sec:superalgebra} we  saw that the only nontrivial interaction between a gauge group and the supersymmetry generators for $N=1$ supersymmetry is given by  (\ref{eq:QR_commutator_dirac}), which in Weyl spinor notation for the supercharges is
\begin{equation}
[Q_A, R] = Q_A, \quad  [\bar Q_{\dot A}, R ] = -\bar Q_{\dot A},
\end{equation}
where $R$ is the generator for an abelian $U(1)$ gauge group.
Using this generator we can define a $U(1)_R$ transformation on the superspace coordinates called {\bf $\mathbf R$-symmetry} as
\begin{equation}
\theta_A \to \theta_A' = e^{i\alpha} \theta_A, \quad \bar\theta_{\dot A} \to  \bar\theta'_{\dot A}= e^{-i\alpha} \bar\theta_{\dot A},
\end{equation}
where $\alpha$ is the parameter of the transformation and the charge of $\theta$ and $\bar\theta$ under the transformation is 1 and $-1$, respectively. It follows from the differential representation of the supersymmetry generators (\ref{eq:diffrepQ}) and (\ref{eq:diffrepQbar}) that we must have
\begin{equation}
Q_A \to Q_A' = e^{-i\alpha} Q_A, \quad \bar Q_{\dot A} \to  \bar Q'_{\dot A}= e^{i\alpha} \bar Q_{\dot A}.
\end{equation}

If a supersymmetric Lagrangian is to be invariant under such a transformation at all then a superfield $\Phi$ must transform as
\begin{equation}
\Phi(x,\theta,\bar\theta)\to \Phi'(x,\theta',\bar\theta')= e^{ir_\Phi\alpha}\Phi(x,\theta, \bar\theta),
\end{equation}
where $r_\Phi$ is the charge of that superfield under the transformation. The charge of a product of two superfields is then just the sum of the charges of the fields. This is required so that the kinetic term in the Lagrangian is invariant under the transformation.

This means that for a scalar (left- or right-handed) superfield the component fields $A$, $\psi$, and $F$ must transform as
\begin{equation}
A\to e^{ir_\Phi\alpha}A,\quad \psi\to e^{i(r_\Phi-1)\alpha} \psi,\quad F\to e^{i(r_\Phi-2)\alpha} F,
\end{equation}
while vector superfields must have zero $R$-charge since they are real, and their component fields, in the Wess--Zumino gauge must thus transform as
\begin{equation}
V_\mu \to V_\mu ,\quad \lambda \to e^{i\alpha} \lambda,\quad D\to D.
\end{equation}
We can check that all the terms in the Lagrangian (\ref{eq:SUSY_Lagrangian}) are invariant under this transformation, with the exception of the superpotential $W$, which must have total $R$-charge 2 for the superpotential term to be invariant. For example, with a single superfield $\Phi$, in the superpotential $W(\Phi)=t\Phi+m\Phi^2+\lambda\Phi^3$, the tadpole term will only be allowed if the $R$-charge of $\Phi$ is 2, the mass term only if the charge is 1 and the Yukawa term only if the charge is $2/3$.
This will exclude most allowed charge assignments for the scalar superfields, and in the minimal supersymmetric models we shall look at there will be no viable continuous $R$-symmetry. However, $R$-symmetry has importance in building models of supersymmetry breaking, and could remain as a broken symmetry in the model.

### 2025-10-20 poincare.tex
Fixes for lecture 21.10.25

Because of (\ref{eq:QQ}) this new algebra is a non-trivial extension of the Poincaré algebra that avoids the no-go theorem. This extension can be proven, under some reasonable assumptions, to be the {\it largest possible} extension of the symmetries of Special Relativity. However, in the $Q_a$ we have introduced new operators that (disappointingly) do not correspond to the generators of the gauge groups, which should in any case  be related by commutators, not anti-commutators. The gauge group generators {\it can} appear in the algebra if we instead extend the algebra with $N>1$ sets of new spinors $Q_a^\alpha$, where $\alpha = 1,\ldots,N$. This gives rise to so-called $N>1$ supersymmetries, while a single set of $Q_a$ is called $N=1$ supersymmetry. Given a gauge group algebra $[B_i,B_j]=iC_{ij}^{~~k}B_k$, we can then extended the superalgebra by the non-trivial commutator $[Q_a^\alpha,B_l]=iS_l^{\alpha\beta}Q_a^\beta$, where $S_l$ are matrix representations of the gauge symmetry group, which does not work for $N=1$. 
% TODO: Extend this to R-symmetry as an abelian alternative. Move some things from later R-symmetry section

>>>>

Because of (\ref{eq:QQ}) this new algebra is a non-trivial extension of the Poincaré algebra that avoids the no-go theorem. This extension can be proven, under some reasonable assumptions, to be the {\it largest possible} extension of the symmetries of Special Relativity. However, in the $Q_a$ we have introduced new operators that (disappointingly) do not correspond to the generators of the gauge groups, which should in any case  be related by commutators, not anti-commutators. The gauge group generators {\it can} appear in the algebra if we instead extend the algebra with $N>1$ sets of new spinors $Q_a^\alpha$, where $\alpha = 1,\ldots,N$. This gives rise to so-called $N>1$ supersymmetries, while a single set of $Q_a$ is called $N=1$ supersymmetry. Given a gauge group algebra $[B_i,B_j]=iC_{ij}^{~~k}B_k$ called the {\bf R-symmetry}, we can then extended the superalgebra by the non-trivial commutator
\begin{equation*}
[Q_a^\alpha,B_l]=S_l^{\alpha\beta}Q_a^\beta,
\end{equation*}
where the $S_l$ are an $N\times N$ matrix representation of the R-symmetry group generators. For $N=1$ supersymmetry this reduces to the abelian
\begin{equation}
[Q_a,B]=Q_a,
\label{eq:QR_commutator_dirac}
\end{equation}
% TODO: Move some things here from later R-symmetry section?

### 2025-10-20 superspace.tex
Fixes for lecture 21.10.25

As a crucial by-product we can now write down a differential representation for the supersymmetry generators by applying the standard expression for the generators $X_i$ of a Lie algebra, given the composition functions $f_\pi$:
\[ iX_j = \frac{\partial f_\pi}{\partial a_j}\frac{\partial}{\partial z_\pi},\]
which gives us:
\begin{eqnarray}
iP_\mu &=& \partial_\mu \label{eq:diffrepP},\\
iQ_A &=& -i(\sigma^\mu\bar{\theta})_A\partial_\mu +\partial_A,\\
i\bar{Q}^{\dot{A}} &=& -i(\bar{\sigma}^\mu \theta)^{\dot{A}}\partial_\mu +\bar\partial^{\dot{A}}.
\label{eq:diffrepQbar}
\end{eqnarray}
The interested reader can now use these expressions to check the (anti-)commutation relations for the supercharges in Eqs.~(\ref{eq:QQweyl}) and (\ref{eq:QQbarweyl}).

@@

Products of the same type of superfield is a superfield of the same type since for left-handed scalar superfields $\Phi_i$ and $\Phi_j$,
\[ \bar{D}_{\dot{A}}(\Phi_i\Phi_j) = (\bar{D}_{\dot{A}}\Phi_i)\Phi_j +\Phi_i( \bar{D}_{\dot{A}}\Phi_j)=0,\]
and similarly for a right-handed scalar superfields, and for vector superfields $V_i$ and $Vi_j$,
\[(V_iV_j)^\dagger=V_j^\dagger V_i^\dagger=V_j V_i= V_iV_j.\]
The product of a left-handed scalar superfield $\Phi$ and its hermitian conjugate $\Phi^\dagger$, $V=\Phi \Phi^\dagger$, is a vector superfield since
\[V^\dagger=(\Phi \Phi^\dagger)^\dagger=\Phi \Phi^\dagger=V.\]
The same is true for sums of superfields of the same type. These properties will be important when creating a superfield version of a Lagrangian.
% TODO: Extend this to the idea of holonomic functions

>>>>

As a crucial by-product we can now write down a differential representation for the supersymmetry generators by applying the standard expression for the generators $X_i$ of a Lie algebra, given the composition functions $f_\pi$:
\[ iX_j = \frac{\partial f_\pi}{\partial a_j}\frac{\partial}{\partial z_\pi},\]
which gives us:
\begin{eqnarray}
iP_\mu &=& \partial_\mu \label{eq:diffrepP},\\
iQ_A &=& -i(\sigma^\mu\bar{\theta})_A\partial_\mu +\partial_A,\label{eq:diffrepQ} \\
i\bar{Q}^{\dot{A}} &=& -i(\bar{\sigma}^\mu \theta)^{\dot{A}}\partial_\mu +\bar\partial^{\dot{A}}.
\label{eq:diffrepQbar}
\end{eqnarray}
The interested reader can now use these expressions to check the (anti-)commutation relations for the supercharges in Eqs.~(\ref{eq:QQweyl}) and (\ref{eq:QQbarweyl}).

@@

Products of the same type of superfield is a superfield of the same type since for left-handed scalar superfields $\Phi_i$ and $\Phi_j$,
\[ \bar{D}_{\dot{A}}(\Phi_i\Phi_j) = (\bar{D}_{\dot{A}}\Phi_i)\Phi_j +\Phi_i( \bar{D}_{\dot{A}}\Phi_j)=0,\]
and similarly for a right-handed scalar superfields, and for vector superfields $V_i$ and $V_j$,
\[(V_iV_j)^\dagger=V_j^\dagger V_i^\dagger=V_j V_i= V_iV_j.\]
The product of a left-handed scalar superfield $\Phi$ and its hermitian conjugate $\Phi^\dagger$, $V=\Phi \Phi^\dagger$, is a vector superfield since
\[V^\dagger=(\Phi \Phi^\dagger)^\dagger=\Phi \Phi^\dagger=V.\]
The same is true for sums of superfields of the same type. These properties will be important when creating a superfield version of a Lagrangian.
% TODO: Extend this to the idea of holonomic functions
